Mining Interaction Log Data in a Creative Arts Practice using Macro Gestures and Transition Matrices

### 2015-09-21 flux-entropy-paper/flux-entropy-paper.tex
updates to conclusion

In this note we have extended previous work on transition
matrices with two simple matrix measures, flux and entropy, both of
which have been demonstrated to have significant discriminatory power
with regards to a corpus of 95 collaborative touch-screen
interactions. Not only do these measures differentiate between
different types of sessions and the internal structure of these
sessions, but they have explanatory power that has thrown new light on
the musical interactions under examination. HCI researchers are
increasingly concerned with interactions that happen in-the-wild,
where the interactions of many users may be collected without the
opportunity to survey or interview users. We suggest that analysing
interaction-state transitions using matrices and simple measures can
lead to new insights into HCI interactions outside the lab.

%TODO Mention Manovich transmutation stuff
The macro-gestures that these processes reveal could also be used as
the basis for new works of new media art~\cite{Manovich:2002ly}.

% Additionally, a wider dissemination of such log data will better
% support the replication of findings in HCI
% research~\cite{Wilson:2011ve}

>>>>

In this note we have extended previous work using transition matrices
to understand HCI processes with the application of two simple matrix
measures, flux and entropy, both of which have been demonstrated to
have significant discriminatory power with regards to a corpus of 95
collaborative touch-screen interactions. Not only do these measures
differentiate between different types of sessions and the internal
structure of these sessions, but they have explanatory power that has
thrown new light on the musical interactions under examination. It is
notable that, for improvised performances, live performances had the
lowest measure in both flux and entropy, while lab-studies had the
highest. This may be due to the increased pressure on performers when
improvising in front of an audience. The internal investigation of
performances session revealed more structural variety in
improvisations than compositions. In response, new works could be
composed that specifically explore different gestural structures in
touch-screen performance

HCI researchers are increasingly concerned with interactions that
happen in-the-wild, where the creative collaborations of many users
may be recorded, but where conducting surveys or interviews would be
innapropriate. We suggest that analysing interaction-state transitions
using matrices and simple measures can lead to new insights into HCI
interactions outside the lab.

Data-mining logs of interactions with creative interfaces sugggest
further approaches for future work. The macro-gestures that these
processes reveal could be used as the basis for new works of new media
art~\cite{Manovich:2002ly}. Additionally, a wider dissemination of
such log data could better support the replication of findings in HCI
research~\cite{Wilson:2011ve}
